---
title: "reporting results from land use model LUISA-BEES"
author: "Sarah"
date: "Apil 1 2020"
output:
  word_document: 
    toc: yes
  html_document: 
      toc: yes
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..") # file path relative to the root of the project directory
#install.packages("E:\\mubarsa\\Documents\\Downloads\\rlang_0.3.1.gz", repos = NULL, type = "source")
#install.packages("ggplot2", dep = TRUE)
#install.packages("dplyr", dep = TRUE)
#install.packages("tidyverse", dep = TRUE)
library(ggplot2)
library(rlang)
library(purrr)
library(dplyr)
library(hms)
library(tidyverse)
library(readxl)
#library(ggplot2)
library(zoo)
library(data.table)
library(foreign)
library(networkD3)
library(rmarkdown)
```


## Create a function that reads xTab files
```{r}
datadir<-"D:\\landuse\\luisetta_outmarch312020\\"
countries<-list.dirs(datadir, full.names = FALSE, recursive=FALSE)
filename<-"P2050_xTabMT.dbf"
#loop through folders to get xTab2050 data
for (i in seq_along(countries)) {
  xTab<-read.dbf(paste0(datadir/datadir[i],filename))

  
}
```
# Graphs of land use change
##Create Sankey diagrams
lu<-unique(as.character(colnames(xTab)))
#create nodes
nodes <- data.frame(node = c(0:17), name = c(lu, "agriculture", "builtup", "forests", "otherNature","newEnergyCrops")) 
#create links
#CREATE new column with land use type in xTab file:
xTab<-cbind(lu,xTab)
#SUM the columns of land uses that can be aggregated:
df<-xTab
df$builtup<-df$URBAN+df$INDUSTRIAL
df$agriculture<-df$ARABLE+df$PERMANENTC+ df$PASTURES+ df$ABANARABLE+ df$ABANPERMAN+ df$ABANPASTUR
df$forests<-df$FORESTSMAT+ df$TRANSWOODL+ df$FORESTSYOU
df$otherNature<-df$SHVA
df$newEnergyCrops<- df$NEWENERGYC

df<-dplyr::select (df,-c(URBAN, INDUSTRIAL, ARABLE, PERMANENTC, PASTURES, FORESTSMAT, TRANSWOODL, ABANARABLE, ABANPERMAN, ABANPASTUR, FORESTSYOU,NEWENERGYC, SHVA))
#MERGE urban+industry=builtup; arable+PC+abandoned etc =agri; forests Y+M+Woodland summed
# format in prep for sankey diagram
links <- tidyr::gather(df, toLU, ha, -lu)
luc <- merge(links, nodes, by.x = "lu", by.y = "name") 
luc <- merge(luc, nodes, by.x = "toLU", by.y = "name")
forSankey <- luc[ , c("node.x", "node.y", "ha")]
colnames(forSankey) <- c("source", "target", "value")
#remove rows containing 0 or else the lines will be drawn in the Sankey:
gtzero<- apply(forSankey, 1, function(row) all(row !=0 ))
##Subset as usual
forSankey<-forSankey[gtzero,]


#make Sankey
networkD3::sankeyNetwork(Links = forSankey, Nodes = nodes, 
Source = 'source', 
Target = 'target', 
Value = 'value', 
NodeID = 'name',
units = 'ha')

#FOR nature, agri and builtup sankey:
#SUM the columns of land uses that can be aggregated:
#create nodes
nodes <- data.frame(node = c(0:15), name = c(lu, "agriculture", "builtup", "nature")) 
df<-xTab
df$builtup<-df$URBAN+df$INDUSTRIAL
df$agriculture<-df$ARABLE+df$PERMANENTC+ df$PASTURES+ df$ABANARABLE+ df$ABANPERMAN+ df$ABANPASTUR+df$NEWENERGYC
df$nature<-df$FORESTSMAT+ df$TRANSWOODL+ df$FORESTSYOU+df$SHVA
df<-dplyr::select (df,-c(URBAN, INDUSTRIAL, ARABLE, PERMANENTC, PASTURES, FORESTSMAT, TRANSWOODL, ABANARABLE, ABANPERMAN, ABANPASTUR, FORESTSYOU,NEWENERGYC, SHVA))
#MERGE urban+industry=builtup; arable+PC+abandoned etc =agri; nature= Y+M+Woodland summed
# format in prep for sankey diagram
links <- tidyr::gather(df, toLU, ha, -lu)
luc <- merge(links, nodes, by.x = "lu", by.y = "name") 
luc <- merge(luc, nodes, by.x = "toLU", by.y = "name")
forSankey <- luc[ , c("node.x", "node.y", "ha")]
colnames(forSankey) <- c("source", "target", "value")
#remove rows containing 0 or else the lines will be drawn in the Sankey:
gtzero<- apply(forSankey, 1, function(row) all(row !=0 ))
##Subset as usual
forSankey<-forSankey[gtzero,]


#make Sankey
networkD3::sankeyNetwork(Links = forSankey, Nodes = nodes, 
Source = 'source', 
Target = 'target', 
Value = 'value', 
NodeID = 'name',
units = 'ha')



## Create Chord Network diagram
```{r}
chordNetwork(Data = xTab, width = 500, height = 500, colourScale =c("#FF0000", "#FF00FF", "#FFDD89", "#FF1212", "#00FF89", "#00FF00","#00FF99", "#FFDD89", "#00FF00", "#FF0000", "#FFDD89", "#00FF00","#00FF00"), labels = c("urban", "industry", "arable", "permCrops", "pastures", "forestMature", "woodland","abanArab", "abanPC", "abanPast", "necr", "shva", "forestY"))
```
