---
title: "exploit capri output for land use and ecosystem services modelling"
author: "Sarah, Paul, Erik Oudejans (Object Vision BV)"
date: "December 15, 2021"
output:
  word_document: 
    toc: yes
  html_document: 
      toc: yes
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..") # file path relative to the root of the project directory
library(ggplot2)
library(rlang)
library(purrr)
library(dplyr)
library(hms)
library(tidyverse)
library(readxl)
library(zoo)
library(data.table)
library(vctrs)
library(gdxrrw)

```

## Parameters
```{r}

fldr_GAMS  <- "C:\\GAMS\\37"
fn_CapriRegions <- "C:\\Users\\Cicada\\Documents\\GeoDMS\\luisabees_r\\data\\CapriRegions.csv"
fldr_datapath <- "C:\\Users\\Cicada\\Documents\\GeoDMS\\luisabees_r\\data"                      # input
fldr_nominmax <- "C:\\Users\\Cicada\\Documents\\GeoDMS\\luisabees_r\\nominmax"                  # output nominmax
fldr_minmax <- "C:\\Users\\Cicada\\Documents\\GeoDMS\\luisabees_r\\enddata_july2019"            # output minmax

```

## Initialize GAMS - eval=false
```{r}
igdx(fldr_GAMS)
```

# Read gdx files directly using rgdx.param from the gdxrrw package
## Read one file
```{r}

gdxfiles <- list.files(fldr_datapath, ".gdx", full.names = TRUE)
gdxfiles[1]
```
## Create a function that reads many files
```{r}

#' @param gdxfile character name of a gdx file
#' @return data frame 
readcaprioutput <- function(gdxfile, symName = "dataout",
                            names = c("RALL","EMPTY","COLS","ROWS","Y","VALUE")){
    message("\nReading ", gdxfile)
    capriout <- rgdx.param(gdxfile, symName = symName, names = names)
    # Change factor to character variables
    capriout$RALL <- as.character(capriout$RALL)
    capriout$EMPTY <- as.character(capriout$EMPTY)
    capriout$COLS <- as.character(capriout$COLS)
    capriout$ROWS <- as.character(capriout$ROWS)
    # Change year from factor to a numeric variable
    capriout$Y <- as.numeric(as.character(capriout$Y))
    return(capriout)
}

#uniqueLanduseTypes <- unique(capriout[c("COLS")]);

# Use the function to read one file (takes 6 seconds)
#system.time(res_2_0810mtr_rd_ref <- readcaprioutput(gdxfiles[1]))

# Use the function to read 13 files (should take more than a minute)
caprilist <- lapply(gdxfiles, readcaprioutput)
capriout <- Reduce(rbind, caprilist)


```
```{r}
capriRegions <- read.csv(file = fn_CapriRegions)
```




# Parse capriout data
## Create demand files for land use classes
```{r}
#unwanted labels EU27yr19   #filter(nchar(RALL)==8) %>%

arable<-capriout %>% 
  select(-EMPTY) %>% 
  filter(COLS == "ARAB") %>%
  filter(ROWS == "LEVL") %>%
  dplyr::arrange(RALL, Y)
perm<-capriout %>% 
  select(-EMPTY) %>% 
  filter(COLS == "PERM") %>% 
  filter(ROWS == "LEVL") %>% 
  dplyr::arrange(RALL, Y)
necr<-capriout %>%            # new economic crops
  select(-EMPTY) %>% 
  filter(COLS == "NECR") %>% 
  filter(ROWS == "LEVL") %>% 
  dplyr::arrange(RALL, Y)
grass<-capriout %>% 
  select(-EMPTY) %>% 
  filter(grepl("GRAI|GRAE" ,COLS)) %>%
  filter(ROWS == "LEVL") %>% 
  dplyr::arrange(RALL, Y) %>%
  group_by(RALL,ROWS,Y) %>%
  summarise(amount_1000ha = sum(VALUE)) 
grass<-add_column(grass, COLS="GRASS", .after=1)
heads<-capriout %>% 
  select(-EMPTY) %>% 
  filter(grepl("HEIF|HEIL|BULL|BULF" ,COLS)) %>%
  filter(ROWS == "LEVL") %>%
  dplyr::arrange(RALL, Y) %>%
  group_by(RALL,ROWS,Y) %>%
  summarise(amount_1000heads = sum(VALUE)) 
heads<-add_column(heads, COLS="COWS", .after=1)
```


#function to select and match given landuse to capriRegions
```{r}
#readLandUseClass <- function(COLS_filter, ROWS_filter, ) {
  
#}

cleanLanduseClass <- function(rawLanduseClass, capriRegions) {
  
  unique_years = unique(rawLanduseClass[c("Y")])
  print(length(capriRegions))
  print(length(unique_years))
  landuseClass <- data.frame(matrix(0, nrow(capriRegions)*nrow(unique_years), 3))
  landuseClass[,1] = capriRegions #[,3]
  landuseClass[,1] = slice(landuseClass, rep(1:n(), each = 3))
  landuseClass[,2] = rep(unique_years, (nrow(capriRegions)))
  colnames(landuseClass) <- c("RALL", "Y", colnames(rawLanduseClass)[5])
  
  key_org  <- paste(substr(rawLanduseClass$RALL, 0,4), rawLanduseClass$Y, sep="", collapse=NULL)
  key_dest <- paste(substr(landuseClass$RALL, 0, 4), landuseClass[,2], sep="", collapse=NULL)
  inds = match(key_org, key_dest)
  landuseClass[inds,3] = rawLanduseClass[,5]
  
  colnames(landuseClass) <- c("RALL", "Y", colnames(rawLanduseClass)[5])
  
  return(landuseClass)
}

#test_unique <- as.vector(array(unique(capriout$RALL)))
unique_regions <- unique(capriout[c("RALL")])

cgrass <- cleanLanduseClass(grass, unique_regions)
cnecr  <- cleanLanduseClass(necr, unique_regions)
cperm <- cleanLanduseClass(perm, unique_regions)
carable <- cleanLanduseClass(arable, unique_regions)

```

## GENERATE DEMAND FILES FOR LUISABEES,PARSE DATA BY COUNTRY  
```{r}
#temp<-bind_cols(arable, perm, grass) 
#temp2<-left_join(temp, necr,  by = c("RALL", "Y"))
#demandall<-temp2 %>%
#  select(-ROWS.x, -COLS.x,-RALL1, -COLS1, -ROWS1, -Y1, -COLS2, -RALL2, -ROWS2, -Y2, -COLS.y,-ROWS.y) %>%
#  rename(Region:=RALL, Year:=Y,OtherArableMax:=VALUE.x, PermCropMax:=VALUE1,PastureMax:=amount_1000ha, NewEnergyMax:=VALUE.y)%>%
#  mutate(OtherArableMax=as.integer(OtherArableMax*1000.),PermCropMax=as.integer(PermCropMax*1000.),PastureMax=as.integer(PastureMax*1000.),New#EnergyMax=as.integer(NewEnergyMax*1000.)) %>%
#  mutate_all(funs(replace(., is.na(.), 0)))



demandall<-bind_cols(carable, cperm[,3], cgrass[,3], cnecr[,3])
#demandall<-select(temp, -one_of("COLS...2", "ROWS...3","RALL...6", "COLS...7", "ROWS...8", "Y...9", "COLS...12", "RALL...11", "ROWS...13", "Y...14"))
demandall <- rename(demandall, Region=RALL, Year=Y, OtherArableMax=VALUE, PermCropMax=...4, PastureMax=...5, NewEnergyMax=...6)
demandall <- mutate(demandall, OtherArableMax=as.integer(OtherArableMax*1000.),PermCropMax=as.integer(PermCropMax*1000.),PastureMax=as.integer(PastureMax*1000.),NewEnergyMax=as.integer(NewEnergyMax*1000.))
demandall <-  mutate_all(demandall, funs(replace(., is.na(.), 0)))

```

## expand to annual values (nomin, nomax)
```{r}
cols <- c("Year", "OtherArableMax","PermCropMax","PastureMax","NewEnergyMax")
temptointerpolate<-setDT(demandall)[, .SD[match(2010:2070, Year)], by = Region]
year_range <- (2010:2070)
temp<-temptointerpolate
temp[,2] = rep(year_range,nrow(unique_regions))

#temptointerpolate[, (cols) := lapply(.SD, na.approx, na.rm = FALSE), .SDcols = cols]
#interpolated<-temptointerpolate%>%  
#  mutate(OtherArable=as.integer(OtherArableMax), PermCrop=as.integer(PermCropMax),Pasture=as.integer(PastureMax), #NewEnergy=as.integer(NewEnergyMax)) %>%
#  select(-OtherArableMax) %>%
#  select(-PermCropMax) %>%
#  select(-PastureMax) %>%
#  select(-NewEnergyMax)
#interpolated$Region<-substr(interpolated$Region,0,4) 

```


## Temporal interpolation
```{r}
cols        <- c("Year", "OtherArableMax","PermCropMax","PastureMax", "NewEnergyMax")
temptointerpolate<-setDT(demandall)[, .SD[match(2010:2070, Year)], by = Region] 
temptointerpolate[, (cols) := lapply(.SD, na.approx, na.rm = FALSE), .SDcols = cols, by=Region] # actual interpolation (na.approx)
year_range <- (2010:2070)
temptointerpolate[,2] = rep(year_range,nrow(unique_regions))

# interpolate forwards
temptointerpolate[, (cols) := lapply(.SD, na.locf, na.rm = FALSE), .SDcols = cols, by=Region]

# interpolate backwards
temptointerpolate <- setDT(map_df(temptointerpolate, rev))
temptointerpolate[, (cols) := lapply(.SD, na.locf, na.rm = FALSE), .SDcols = cols, by=Region]
temptointerpolate <- map_df(temptointerpolate, rev)

interpolated<-temptointerpolate%>%  
  mutate(OtherArable=as.integer(OtherArableMax), PermCrop=as.integer(PermCropMax),Pasture=as.integer(PastureMax), NewEnergy=as.integer(NewEnergyMax)) %>%
  select(-OtherArableMax) %>%
  select(-PermCropMax) %>%
  select(-PastureMax) %>%
  select(-NewEnergyMax)
interpolated$Region<-substr(interpolated$Region,0,4)

```
## write to files 
```{r}



dir.create(fldr_nominmax)


interpolated %>% 
  mutate(countryiso2 = substr(Region, 1,2)) %>% 
  group_by(countryiso2) %>% 
  nest() %>% 
  walk2(
    .x = .$countryiso2,
    .y = .$data,
    .f = ~ write_csv(.y, file.path(fldr_nominmax, str_c(.x, ".csv")))
   #.f = ~ write_csv(.y, file.path(enddatanew, str_c(.x, ".csv")))
  )


```

## expand to annual and min values
```{r}
cols <- c("Year", "OtherArableMax","PermCropMax","PastureMax","NewEnergyMax")

temptointerpolate<-setDT(demandall)[, .SD[match(2010:2070, Year)], by = Region] 
temptointerpolate[, (cols) := lapply(.SD, na.approx, na.rm = FALSE), .SDcols = cols, by=Region] # actual interpolation (na.approx)
year_range <- (2010:2070)
temptointerpolate[,2] = rep(year_range,nrow(unique_regions))

# interpolate forwards
temptointerpolate[, (cols) := lapply(.SD, na.locf0), .SDcols = cols, by=Region]

# interpolate backwards
temptointerpolate <- setDT(map_df(temptointerpolate, rev))
temptointerpolate[, (cols) := lapply(.SD, na.locf0), .SDcols = cols, by=Region]
temptointerpolate <- map_df(temptointerpolate, rev)

interpolated<-temptointerpolate %>%
add_column(OtherArableMin = 0, .before = "OtherArableMax") %>%
add_column(PermCropMin = 0, .before = "PermCropMax") %>%
add_column(PastureMin = 0, .before = "PastureMax") %>%
add_column(NewEnergyMin = 0, .before = "NewEnergyMax")

interpolated$Region<-substr(interpolated$Region,0,4)
interpolated$OtherArableMin = as.integer(interpolated$OtherArableMax*0.7)
interpolated$PermCropMin    = as.integer(interpolated$PermCropMax*0.7)
interpolated$PastureMin     = as.integer(interpolated$PastureMax*0.7)
interpolated$NewEnergyMin   = as.integer(interpolated$NewEnergyMax*0.7)

interpolated$OtherArableMax = as.integer(interpolated$OtherArableMax)
interpolated$PermCropMax    = as.integer(interpolated$PermCropMax)
interpolated$PastureMax     = as.integer(interpolated$PastureMax)
interpolated$NewEnergyMax   = as.integer(interpolated$NewEnergyMax)

#interpolated %>% mutate(OtherArableMax=as.integer(OtherArableMax), PermCropMax=as.integer(PermCropMax),PastureMax=as.integer(PastureMax),NewEnergyMax=as.integer(NewEnergyMax)) 

  
```



#countries<-unique(substr(arable$RALL,1,2))
## write to files 
```{r}
dir.create(fldr_minmax)

interpolated %>% 
  mutate(countryiso2 = substr(Region, 1,2)) %>% 
  group_by(countryiso2) %>% 
  nest() %>% 
  walk2(
    .x = .$countryiso2,
    .y = .$data,
   .f = ~ write_csv(.y, file.path(fldr_minmax, str_c(.x, "_new.csv")))
   #.f = ~ write_csv(.y, file.path(enddatanew, str_c(.x, ".csv")))
  )

```


