---
title: "exploit capri output for land use and ecosystem services modelling"
author: "Sarah, Paul"
date: "February 13, 2019"
output:
  word_document: 
    toc: yes
  html_document: 
      toc: yes
---

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "..") # file path relative to the root of the project directory
#install.packages("E:\\mubarsa\\Documents\\Downloads\\rlang_0.3.1.gz", repos = NULL, type = "source")
#install.packages("ggplot2", dep = TRUE)
#install.packages("dplyr", dep = TRUE)
#install.packages("tidyverse", dep = TRUE)
library(ggplot2)
library(rlang)
library(purrr)
library(dplyr)
library(hms)
library(tidyverse)
library(readxl)
#library(ggplot2)
#(gdxrrw)
library(zoo)
library(data.table)

# Setup Sarah's machine
#igdx(gamsSysDir = "C:\\GAMS\\win64\\26.1")
```

# Install packages

## Install the GDXRRW package to read gdx data
```{r eval=FALSE}
#install.packages("C:\\LUISETTA\\Rproject\\gdxrrw_1.0.2.tar.gz", repos = NULL, type = "source")
library(gdxrrw)
igdx()
datapath <- "C:\\LUISETTA\\Rproject\\ReferenceScenarioLTS\\original\\"

```
# Read gdx files directly using rgdx.param from the gdxrrw package
## Read one file
```{r}

gdxfile <- list.files(datapath, ".gdx", full.names = TRUE)
gdxfile[1]
```
## Create a function that reads many files
```{r}

#' @param gdxfile character name of a gdx file
#' @return data frame 
readcaprioutput <- function(gdxfile, symName = "dataout",
                            names = c("RALL","EMPTY","COLS","ROWS","Y","VALUE")){
    message("\nReading ", gdxfile)
    capriout <- rgdx.param(gdxfile, symName = symName, names = names) 
    # Change factor to character variables
    capriout$RALL <- as.character(capriout$RALL)
    capriout$EMPTY <- as.character(capriout$EMPTY)
    capriout$COLS <- as.character(capriout$COLS)
    capriout$ROWS <- as.character(capriout$ROWS)
    # Change year from factor to a numeric variable
    capriout$Y <- as.numeric(as.character(capriout$Y))
    return(capriout)
}

# Use the function to read one file (takes 6 seconds)
#system.time(res_2_0810mtr_rd_ref <- readcaprioutput(gdxfiles[1]))

# Use the function to read 13 files (should take more than a minute)
caprilist <- lapply(gdxfiles, readcaprioutput)
capriout <- Reduce(rbind, caprilist)


# save the output to an R file
saveRDS(capriout, file.path(datapath, "capriout.rds"))

if(FALSE){ # Read the file
    readRDS(file.path(datapath, "capriout.rds"))
}
```

# Parse capriout data
## Create demand files for land use classes
```{r}
arable<-capriout %>% 
  select(-EMPTY) %>% 
  filter(COLS == "ARAB") %>%
  filter(ROWS == "LEVL") %>% 
  dplyr::arrange(RALL, Y)  
perm<-capriout %>% 
  select(-EMPTY) %>% 
  filter(COLS == "PERM") %>% 
  filter(ROWS == "LEVL") %>% 
  dplyr::arrange(RALL, Y)
necr<-capriout %>% 
  select(-EMPTY) %>% 
  filter(COLS == "NECR") %>% 
  filter(ROWS == "LEVL") %>% 
  dplyr::arrange(RALL, Y)
grass<-capriout %>% 
  select(-EMPTY) %>% 
  filter(grepl("GRAI|GRAE" ,COLS)) %>%
  filter(ROWS == "LEVL") %>% 
  dplyr::arrange(RALL, Y) %>%
  group_by(RALL,ROWS,Y) %>%
  summarise(amount_1000ha = sum(VALUE)) 
grass<-add_column(grass, COLS="GRASS", .after=1)
heads<-capriout %>% 
  select(-EMPTY) %>% 
  filter(grepl("HEIF|HEIL|BULL|BULF" ,COLS)) %>%
  filter(ROWS == "LEVL") %>% 
  dplyr::arrange(RALL, Y) %>%
  group_by(RALL,ROWS,Y) %>%
  summarise(amount_1000heads = sum(VALUE)) 
heads<-add_column(heads, COLS="COWS", .after=1)

```
# Parse forest data
## Create demand files
```{r}
rawforest <- read_excel("D:\\landuse\\LTS_rawdata\\IIASA_Reference2016_2070_forest_area_G4M_iso.xlsx") 
longforest <-rawforest %>%
  gather(year,forestarea,as.character(2000):as.character(2070)) %>%
  arrange(iso) 
longforestnoblcf<-longforest[-grep("areaused", longforest$attribute),]
forestLTS<-ddply(longforestnoblcf,c("iso", "year"), summarise, forestarea=sum(forestarea)) 
demandforest<-forestLTS %>%
  rename(Region:=X__1, Year:=year,Forest:=forestarea)%>%
  mutate_all(funs(replace(., is.na(.), 0)))

```
## expand forest data to annual values (nomin, nomax)
```{r}
cols <- c("Year", "Forest")
#requires library(data.table)
temptointerpolate <- setDT(demandforest)[, .SD[match(2000:2070, Year)], by = Region] 
#requires library(zoo)
temptointerpolate[, (cols) := lapply(.SD, na.approx, na.rm = FALSE), .SDcols = cols]
forestinterpolated <- temptointerpolate %>%  
  mutate(ForestsMatMin=as.integer(Forest*0.7),ForestsMatMax=as.integer(Forest*1.3), ForestsYoMin = 0,ForestsYoMax=0)

```
## write to files 
```{r}

forestLTSfinal <- "E:\\data\\globiomg4m\\demandluisabees\\"
#requires library(purrr)
forestinterpolated %>% 
  mutate(countryiso2 = substr(Region, 1,2)) %>% 
  group_by(countryiso2) %>% 
  nest() %>% 
  walk2(
    .x = .$countryiso2,
    .y = .$data,
    .f = ~ write.csv(.y, file.path(enddata, str_c(.x, ".csv")), row.names=FALSE)
  )

```

## GENERATE DEMAND FILES FOR LUISABEES,PARSE DATA BY COUNTRY  
## isolate useful data
```{r}
temp<-bind_cols(arable, perm, grass) 
temp2<-left_join(temp, necr,  by = c("RALL", "Y"))
demandall<-temp2 %>%
  select(-ROWS.x, -COLS.x,-RALL1, -COLS1, -ROWS1, -Y1, -COLS2, -RALL2, -ROWS2, -Y2, -COLS.y,-ROWS.y) %>%
  rename(Region:=RALL, Year:=Y,OtherArableMax:=VALUE.x, PermCropMax:=VALUE1,PastureMax:=amount_1000ha, NewEnergyMax:=VALUE.y)%>%
  mutate(OtherArableMax=as.integer(OtherArableMax*1000.),PermCropMax=as.integer(PermCropMax*1000.),PastureMax=as.integer(PastureMax*1000.),NewEnergyMax=as.integer(NewEnergyMax*1000.)) %>%
  mutate_all(funs(replace(., is.na(.), 0)))

```

## expand to annual values (nomin, nomax)
```{r}
cols <- c("Year", "OtherArableMax","PermCropMax","PastureMax","NewEnergyMax")
temptointerpolate<-setDT(demandall)[, .SD[match(2010:2070, Year)], by = Region] 
e
interpolated<-temptointerpolate%>%  
  mutate(OtherArable=as.integer(OtherArableMax), PermCrop=as.integer(PermCropMax),Pasture=as.integer(PastureMax), NewEnergy=as.integer(NewEnergyMax)) %>%
  select(-OtherArableMax) %>%
  select(-PermCropMax) %>%
  select(-PastureMax) %>%
  select(-NewEnergyMax)
interpolated$Region<-substr(interpolated$Region,0,4) 

```
## write to files 
```{r}

enddata <- "C:\\LUISETTA\\Rproject\\capri4luisabees\\enddata_july2019\\nominmax"

interpolated 

interpolated %>% 
  mutate(countryiso2 = substr(Region, 1,2)) %>% 
  group_by(countryiso2) %>% 
  nest() %>% 
  walk2(
    .x = .$countryiso2,
    .y = .$data,
    .f = ~ write_csv(.y, file.path(enddata, str_c(.x, ".csv")))
   #.f = ~ write_csv(.y, file.path(enddatanew, str_c(.x, ".csv")))
  )

```
## expand to annual and min values
```{r}
cols <- c("Year", "OtherArableMax","PermCropMax","PastureMax","NewEnergyMax")
temptointerpolate<-setDT(demandall)[, .SD[match(2010:2070, Year)], by = Region] 
temptointerpolate[, (cols) := lapply(.SD, na.approx, na.rm = FALSE), .SDcols = cols]
interpolated<-temptointerpolate%>%  
  add_column(OtherArableMin = 0, .before = "OtherArableMax")%>%
  add_column(PermCropMin = 0, .before = "PermCropMax")%>%
  add_column(PastureMin = 0, .before = "PastureMax")%>%
  add_column(NewEnergyMin = 0, .before = "NewEnergyMax") %>%
  mutate(OtherArableMin=as.integer(OtherArableMax*0.7), PermCropMin=as.integer(PermCropMax*0.7),PastureMin=as.integer(PastureMax*0.7), NewEnergyMin=as.integer(NewEnergyMax*0.7)) %>%
  mutate(OtherArableMax=as.integer(OtherArableMax), PermCropMax=as.integer(PermCropMax),PastureMax=as.integer(PastureMax), NewEnergyMax=as.integer(NewEnergyMax)) 
interpolated$Region<-substr(interpolated$Region,0,4)
  
```
#countries<-unique(substr(arable$RALL,1,2))
## write to files 
```{r}

enddata <- "C:\\LUISETTA\\Rproject\\capri4luisabees\\enddata_july2019"
#enddatanew <- "E:/dev/CAPRI/capri4luisabees/enddatanew/"

interpolated 

interpolated %>% 
  mutate(countryiso2 = substr(Region, 1,2)) %>% 
  group_by(countryiso2) %>% 
  nest() %>% 
  walk2(
    .x = .$countryiso2,
    .y = .$data,
   .f = ~ write_csv(.y, file.path(enddata, str_c(.x, "_new.csv")))
   #.f = ~ write_csv(.y, file.path(enddatanew, str_c(.x, ".csv")))
  )

```
## PLOT LTS DATA AGAINST RefScen2017 DATA  ; OVERLAY CLC2012 data
## open original data files
```{r}
oldcapripath <- "E:\\dev\\LUISETTA_originaldeliveryOct2018\\Demand\\StatusQuo\\Capri\\"
oldforestpath <- "E:\\dev\\Luisetta\\Demand\\StatusQuo\\ForestDemandRegions\\"
newcapripath  <- "E:\\dev\\CAPRI\\capri4luisabees\\enddata\\"
newforestpath <- "E:\\data\\globiomg4m\\demandluisabees\\"

#Tidy CLC data (gather - wide to long)
# requires libraries tibble, plyr, tidyr
clcdata <- read_excel("E:\\data\\Land_Cover\\LUISA_base_map_\\tabArea_LU_N0.xlsx")
landuses <- c("urban","industry","infrastructure","greenurban","crops", "permcrops", "pasture", "forest",
              "grassland","transwood","novegnatural", "wetland", "water")
clc<-gather(clcdata, VT_CODE) %>%
  add_column(landuse = "0", .before = "value") 
clc$landuse=rep(landuses, 35)
clc$source="clc"
#TEST PLOT
#ggplot(subset(clc, VT_CODE=="AT"),aes(x=landuse, y=value,group=landuse)) + geom_point() +labs (title = "AT",y="hectares", x="landuse")
oldcaprifiles<-list.files(oldcapripath)
newcaprifiles<-list.files(newcapripath)
oldforestfiles<-list.files(oldforestpath)
newforestfiles<-list.files(newforestpath)

for (i in seq_along(oldfiles)) {
  country<-strsplit(oldfiles[i],".csv")[[1]][1]
  print(country)
  oldcapridata<-read.csv(paste0(oldcapripath,oldcaprifiles[i])) 
  oldcapridata$source<-"rs17"
  oldforestdata<-read.csv(paste0(oldforestpath,oldcaprifiles[i])) 
  oldforestdata$source<-"rs17"
  newcapridata<-read.csv(paste0(newcapripath,oldcaprifiles[i])) 
  newcapridata$source<-"lts"
  newforestdata<-read.csv(paste0(newforestpath,oldcaprifiles[i])) 
  newforestdata$source<-"lts"
  clcloop<-subset(clc, VT_CODE ==country)
  clccrops    <- subset(clcloop, landuse == "crops")
  clcpasture  <- subset(clcloop, landuse == "pasture")
  clcgrass    <- subset(clcloop, landuse == "grassland")
  clcpermcrop <- subset(clcloop, landuse == "permcrops")
  clcforest   <- subset(clcloop, landuse == "forest")
  #column bind OLD capri and OLD forest data. Get rid of trailing "00" on country isocode
  n0<-grep("00", oldcapridata$Region)
  oldcapridata<-oldcapridata[n0,]
  joinolddata<-left_join(oldcapridata,oldforestdata, by = c("Year")) %>%
    mutate(source = source.y) %>%
    add_column(Forest = 0, .before = "ForestsMatMin") %>%
    mutate(Forest=ForestsMatMax + ForestsYoMax) %>%
    select(-source.y, -source.x, -Region.y)
  names(joinolddata)[1]<-"Region"
  #column bind NEW capri and NEW forest data. Get rid of trailing "00" on country isocode
  n0<-grep("00", newcapridata$Region)
  newcapridata<-newcapridata[n0,]
  joinnewdata<-left_join(newcapridata,newforestdata, by = c("Year")) %>%
    mutate(source = source.y) %>%
    select(-source.x, -source.y, -Region.y)
  names(joinnewdata)[1]<-"Region"
 #####
   #stack all new and old to make graphs
  forplot<-rbind(joinolddata, joinnewdata)
#  forplot<-rbind(oldcapridata, newcapridata)
  n0<-grep("00",forplot$Region)
  necr<-forplot[n0,]$NewEnergyMax
  plot<-ggplot(forplot[n0,],aes(y=PastureMax, x=Year, group=source, colour="pasturemax")) +   geom_point(aes(shape=source)) + labs(title=country, y="hectares", x = "year", colour = "landuse")  +  
#add forest
geom_point(aes(y=Forest, color="forest",shape=source) ) + 
#add crops
geom_point(aes(y=OtherArableMax, color="arablemax",shape=source) )+ geom_point(aes(y=PermCropMax, color="permcropmax",shape=source) )+ geom_point(aes(y=NewEnergyMax, color="necrmax",shape=source) ) +  geom_hline(yintercept = (clcpasture$value), color = "grey", linetype = "dashed", size = 1.5) +  geom_hline(yintercept = (clcpasture$value+clcgrass$value), color = "cornflowerblue", linetype = "dashed", size = 1.5) + geom_hline(yintercept = clccrops$value, color = "red", linetype = "dashed", size = 1.5) + geom_hline(yintercept = clcpermcrop$value, color = "purple", linetype = "dashed", size = 1.5) + geom_hline(yintercept = clcforest$value, color = "darkkhaki", linetype = "dashed", size = 1.5)+ labs(subtitle = "dashed lines represent CLC values in 2012:\n green=forest; red=arable; permcrops=purple; teal=pasture + grassland; grey=pasture only")
 
  
  ggsave(paste0("E:\\dev\\CAPRI\\capri4luisabees\\graphs\\",country,"forestsmax.pdf"))
 
#   plotmin<-ggplot(forplot[n0,],aes(y=PastureMin, x=Year, group=source, colour="pasturemin")) + #geom_point(aes(shape=source)) +labs(title=country, y="hectares", x = "year", colour = "landuse")  + #geom_point(aes(y=OtherArableMin, color="arablemin",shape=source) )+ geom_point(aes(y=NewEnergyMin, #color="necrmin",shape=source) )+geom_hline(yintercept = clcpasture$value, color = "blue", linetype = "dashed", size = #1.5) +geom_hline(yintercept = clccrops$value, color = "red", linetype = "dashed", size = 1.5) + labs(subtitle = #"dashed lines represent CLC values in 2012")
#  ggsave(paste0("C:\\LUISETTA\\Rproject\\results\\",country,"min.pdf"))

  }
```


## Create fertilizer files for land use classes
```{r}
arablef<-capriout %>% 
  select(-EMPTY) %>% 
  filter(COLS == "ARAB") %>% 
  filter(grepl("NITF|PHOF|POTF" ,ROWS)) %>%
  dplyr::arrange(RALL, Y) %>% 
  group_by(RALL, COLS, Y) %>% 
  summarise(fert_kg_ha = sum(VALUE))
write.csv(arablef, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/arable_fert.csv")
permf<-capriout %>% 
  select(-EMPTY) %>% 
  filter(COLS == "PERM") %>% 
  filter(grepl("NITF|PHOF|POTF" ,ROWS)) %>%
  dplyr::arrange(RALL, Y) %>% 
  group_by(RALL, COLS, Y) %>% 
  summarise(fert_kg_ha = sum(VALUE))
write.csv(permf, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/perm_fert.csv")
necrf<-capriout %>% 
  select(-EMPTY) %>% 
  filter(COLS == "NECR") %>% 
  filter(grepl("NITF|PHOF|POTF" ,ROWS)) %>%
  dplyr::arrange(RALL, Y) %>% 
  group_by(RALL, COLS, Y) %>% 
  summarise(fert_kg_ha = sum(VALUE))
write.csv(necrf, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/necr_fert.csv")
grassf<-capriout %>% 
  select(-EMPTY) %>% 
  group_by(RALL, COLS, Y) %>% 
  filter(grepl("GRAI|GRAE" ,COLS)) %>%
  filter(grepl("NITF|PHOF|POTF" ,ROWS)) %>%
  dplyr::arrange(RALL, Y) %>% 
  group_by(RALL, Y) %>% 
  summarise(fert_kg_ha = sum(VALUE)) %>%
grassf<-add_column(grassf, COLS="GRASS", .after=1)
write.csv(grassf, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/grass_fert.csv")

```
#*****************************************
# IF USING EXCEL OUTPUT FROM GDX READER:
##Read and format all CAPMOD LEVL sheets
```{r}
pathexcellevl <- file.path(datapath, "CAPRI_lts_RefScenoutput_parsed_levl.xlsx")

# Read all sheets, keep sheet names as a variable and store data in a list-column
levl <- data_frame(sheet = excel_sheets(pathexcellevl)) %>% 
  # Load data from all sheets
  mutate(data = map2(pathexcellevl, sheet, readxl::read_excel, range =  "A1:F9999")) %>% 
  unnest() %>% 
  filter(!is.na(region))
```
```{r}
levl$year<-(as.character(levl$year))
levl$amount_1000ha<-(as.integer(levl$amount_1000ha))

```
# Create demand files for land use classes
```{r}
arable<-levl %>% 
  select(-empty, -attribute,-sheet) %>% 
  filter(landuse == "ARAB") %>% 
  dplyr::arrange(region)  
write.csv(arable, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/arable.csv")

perm<-levl %>% 
  select(-empty, -attribute,-sheet) %>% 
  filter(landuse == "PERM") %>% 
  dplyr::arrange(region)  
write.csv(perm, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/perm.csv")

necr<-levl %>% 
  select(-empty, -attribute,-sheet) %>% 
  filter(landuse == "NECR") %>% 
  dplyr::arrange(region)  
write.csv(necr, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/necr.csv")

grass<-levl %>% 
  select(-empty, -attribute,-sheet) %>% 
  dplyr::arrange(region)  %>%
  group_by(region, year, landuse) %>% 
  filter(grepl("GRAI|GRAE" ,landuse)) %>%
  group_by(region,year) %>%
  summarise(amount_1000ha = sum(amount_1000ha)) 
grass<-add_column(grass, landuse="GRASS", .after=1)
write.csv(grass, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/grass.csv")

livestock<-levl %>% 
  select(-empty, -attribute,-sheet) %>% 
  filter(grepl("BUL|HEI" ,landuse))%>% 
  dplyr::arrange(region)  %>% 
  group_by(region, year) %>% 
  summarise(numheads_1000s = sum(amount_1000ha))
write.csv(livestock, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/livestock.csv")
```

```{r}
# Read all sheets the purr way, keep sheet names as a variable and store data in a list-column
pathexcelfert <- file.path(datapath, "CAPRI_lts_RefScenoutput_parsed_fert.xlsx")
# Read all sheets the purr way, keep sheet names as a variable and store data in a list-column
fert <- data_frame(yr = excel_sheets(pathexcelfert)) %>%
  # Load data from all sheets
   mutate(data = map2(pathexcelfert, yr, readxl::read_excel, range =  "A1:F9999")) %>% 
  unnest() %>% 
  filter(!is.na(region))
```
```{r}
fert$year<-(as.character(fert$year))
fert$amount_kgperha<-(as.integer(fert$amount_kgperha))
```
# Create fert intensity files for land use classes
```{r}
arablef<-fert %>% 
  select(-empty, -yr) %>% 
  filter(landuse == "ARAB") %>% 
  dplyr::arrange(region)  %>%
  group_by(region, year, landuse) %>% 
  filter(!grepl("FERT" ,fertilizer)) %>% 
  summarise(fert_kg_ha = sum(amount_kgperha))
write.csv(arablef, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/arable_fert.csv")

permf<-fert %>% 
  select(-empty, -yr) %>% 
  filter(landuse == "PERM") %>% 
  dplyr::arrange(region)  %>%
  group_by(region, year, landuse) %>% 
  filter(!grepl("FERT" ,fertilizer)) %>% 
  summarise(fert_kg_ha = sum(amount_kgperha))
write.csv(permf, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/perm_fert.csv")

necrf<-fert %>% 
  select(-empty, -yr) %>% 
  filter(landuse == "NECR") %>% 
  dplyr::arrange(region)  %>%
  group_by(region, year, landuse) %>% 
  filter(!grepl("FERT" ,fertilizer)) %>% 
  summarise(fert_kg_ha = sum(amount_kgperha))
write.csv(necrf, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/necr_fert.csv")

grassf<-fert %>% 
  select(-empty, -yr) %>% 
  #summarise(GRASS = sum(amount_kgperha)) %>%
  dplyr::arrange(region)  %>%
  group_by(region, year, landuse) %>% 
  filter(!grepl("FERT" ,fertilizer)) %>% 
  filter(grepl("GRAI|GRAE" ,landuse)) %>%
  summarise(fert_kg_ha = sum(amount_kgperha)) %>%
    summarise(fert_kg_ha = sum(fert_kg_ha)) 
write.csv(grassf, file="C:/Users/mubarsa/jrcbox/capri/ReferenceScenarioLTS/grass_fert.csv")
#************************





S2BIOMcoefs <- data_frame(countryiso = excel_sheets(pathexcel)) %>% 
    # Load data from all sheets
    mutate(data = map2(pathexcel, countryiso, readxl::read_excel, range =  "V4:AD7")) %>% 
  unnest() %>% 
  dplyr::rename(var = X__1) %>% 
  left_join(data_frame(var = c("moisture, % ar",
                               "bulk density, kg/m3 ar",
                               "basic density, kg/m3 ar"),
                       shortvar = c("moisture", "bulkdensity", "basicdensity")),
            by = "var")

 
S2BIOMcoefs <- S2BIOMcoefs %>% 
  gather(product, value, -var, -countryiso,-shortvar) %>%
  select(-var) %>%
  spread(shortvar, value)

S2BIOM <- S2BIOMktdm %>%
  left_join(S2BIOMeurtdm, by = c("countryiso","nuts3","product")) %>%
  left_join(S2BIOMcoefs, by = c("countryiso","product"))
          
```

#Convert units to be compatible with FUSION/CBM
```{r}
S2BIOM$cubicm<- ((S2BIOM$mass*10^6)/S2BIOM$basicdensity )/(1-(S2BIOM$moisture/100.))
S2BIOM$eurpercubicm<- (S2BIOM$cost/10^3)*(S2BIOM$basicdensity/(1-(S2BIOM$moisture/100.))) 
```
#Create new file with ordered min/max costs by 2 categories: "stemwood and thinnings" and "primary residues" 
    
```{r}
#categorise product into sawlogs and residues
S2BIOM$cat<-"Un"
S2BIOM[grep("111", S2BIOM$product),]$cat<-"stemwood&thinnings"
S2BIOM[grep("121", S2BIOM$product),]$cat<-"primaryresidues"
```

#Perform cumulative sum operation based on product
```{r}
# #test to be sure it is working (using raw data)
# testondrytonsw<-S2BIOM %>% 
#   filter(cat == "stemwood&thinnings") %>%
#   filter(countryiso == "NL") %>% 
#   select(-basicdensity,-bulkdensity,-moisture) %>% 
#   arrange(countryiso,cost)  %>% 
#   dplyr::mutate(cumulvol = cumsum(mass))
# testondrytonpr<-S2BIOM %>% 
#   filter(cat == "primaryresidues") %>%
#   filter(countryiso == "NL") %>% 
#   select(-basicdensity,-bulkdensity,-moisture) %>% 
#   arrange(countryiso,cost)  %>% 
#   dplyr::mutate(cumulvol = cumsum(mass))
# test4plot<-rbind(testondrytonsw,testondrytonpr)
# #END TEST

#primary residues
pr<-S2BIOM %>% 
  select(-mass,-cost,-basicdensity,-bulkdensity,-moisture) %>% 
  filter(cat == "primaryresidues") %>% 
  dplyr::arrange(countryiso,eurpercubicm)  %>% 
  dplyr::group_by(countryiso) %>% 
  dplyr::mutate(cumulvol = cumsum(cubicm)) 

#stemwood
sw<-S2BIOM %>% 
  select(-mass,-cost,-basicdensity,-bulkdensity,-moisture) %>% 
  filter(cat == "stemwood&thinnings") %>% 
  arrange(countryiso,eurpercubicm)  %>% 
  dplyr::group_by(countryiso) %>% 
  dplyr::mutate(cumulvol = cumsum(cubicm))
#bind rows together
S2BIOM4plot<-rbind(sw,pr)
```
<!-- #Test on one country to be sure data is correct: -->
<!-- ```{r} -->
<!-- plot<-ggplot(test4plot, -->
<!--              aes(x=cost, y=cumulvol, colour=cat)) + geom_point() + -->
<!--              labs(title="NL TEST", -->
<!--              y="tons dry matter", x = "eur/tdm", colour = "Category") -->
<!-- ggsave("TESTNL_plot.pdf") -->
<!-- ``` -->

#Faceted graph results (too crammed)
# ```{r} 
# pf<- ggplot(S2BIOM4plot, aes(x=eurpercubicm, y=cumulvol, colour=cat)) + geom_point() 
# pf + facet_grid(rows = vars(countryiso), cols = vars(cat))
# pc<- ggplot(S2BIOM4plot, aes(x=eurpercubicm, y=cumulvol, colour=cat)) 
# pc + geom_point(aes(group = countryiso))pr<-S2BIOM %>% 
#   select(-mass,-cost,-basicdensity,-bulkdensity,-moisture) %>% 
#   filter(cat == "PR") %>% 
#   arrange(countryiso,eurpercubicm)  %>% 
#   dplyr::group_by(countryiso) %>% 
#   dplyr::mutate(cumulvol = cumsum(cubicm)) %>% 
# ```
#One graph per country
```{r} 
countrylist<-unique(S2BIOM4plot$countryiso) 
for (i in seq_along(countrylist)) {
plot<-ggplot(subset(S2BIOM4plot, S2BIOM4plot$countryiso == countrylist[i]),
             aes(y=eurpercubicm, x=cumulvol, colour=cat)) + geom_point() +
             labs(title=countrylist[i],
             x="cubic meters", y = "eur/cubic meter", colour = "Category") 
ggsave(paste(countrylist[i],"EuronY_plot.pdf")) 
}
#get cumulative volume for RW and FW for each country and save to variable:
RWvolbycountry<-S2BIOM4plot %>%
    group_by(countryiso) %>%
    filter(cat =="stemwood&thinnings")%>%
    summarise(S2B_RWv = max(cumulvol), S2B_RWmedcost=median(eurpercubicm)) %>%
    mutate(F_RWv='na',diffRWv='na', F_RWmedcost='na', diffRWc='na')

PRvolbycountry<-S2BIOM4plot %>%
    group_by(countryiso) %>%
    filter(cat =="primaryresidues")%>%
    summarise(S2B_PRv = max(cumulvol), S2B_PRmedcost=median(eurpercubicm))%>%
    mutate(F_PRv='na', diffPRv='na', F_PRmedcost='na', diffPRc='na')
```
# GET FUSION OUTPUT
```{r}
rwpath <-paste0(datapath, "/dataFUSION/RW/")
rwfiles<-list.files(rwpath)
prpath <-paste0(datapath, "/dataFUSION/PR/")
prfiles<-list.files(prpath)

for (i in seq_along(rwfiles)) {
  country<-strsplit(rwfiles[i],"_")[[1]][1]
  print(country)
  frw<-read.csv(paste0(rwpath,rwfiles[i])) %>% 
    #subset(supply_row_rel > 0) %>% 
    select(-volume) %>% 
    mutate(cat = "roundwood", countryiso = country)
  fpr<-read.csv(paste0(prpath,prfiles[i])) %>% 
   # subset(supply_row_rel > 0) %>% 
    select(-volume) %>% 
    mutate(cat = "otherwoodcomponents", countryiso = country)
  #stack the two to make graphs
  FUSION4plot<-rbind(frw,fpr)
  plot<-ggplot(FUSION4plot,aes(y=Values, x=cumulvol, colour=cat)) + geom_point() +
             labs(title=country,
             x="cubic meters", y = "eur/cubic meter", colour = "Category") 
  ggsave(paste0(datapath,"/graphsFUSION/",country,"_EuronY_plot.pdf"))
  PRfusion<-FUSION4plot %>%
    filter(cat =="otherwoodcomponents")%>%
    summarise(cumulvol = max(cumulvol))
  print("max cumulative volume PR=")
  print(PRfusion$cumulvol)
  PRfusionc<-FUSION4plot %>%
    filter(cat =="otherwoodcomponents")%>%
    summarise(mediancost = median(Values))
  print("median cost PR=")
  print(PRfusionc$mediancost)
  PRvolbycountry[PRvolbycountry$countryiso == country ,4]<-PRfusion
  PRvolbycountry[PRvolbycountry$countryiso == country ,5]<- round(PRvolbycountry$S2B_PRv - PRfusion)
  PRvolbycountry[PRvolbycountry$countryiso == country ,6]<- PRfusionc
  PRvolbycountry[PRvolbycountry$countryiso == country ,7]<- round(PRvolbycountry$S2B_PRmedcost -  PRfusionc)
  RWfusion<-FUSION4plot %>%
    filter(cat =="roundwood")%>%
    summarise(cumulvol = max(cumulvol))
  print(RWfusion)
  RWfusionc<-FUSION4plot %>%
    filter(cat =="roundwood")%>%
    summarise(mediancost = median(Values))
  print(RWfusionc)
  RWvolbycountry[RWvolbycountry$countryiso == country ,4]<- RWfusion
  #RWvolbycountry[RWvolbycountry$countryiso == country ,5]<- round(RWvolbycountry$S2B_RWv - RWfusion)
  RWvolbycountry[RWvolbycountry$countryiso == country ,6]<- RWfusionc
  #RWvolbycountry[RWvolbycountry$countryiso == country ,7]<- round(RWvolbycountry$S2B_RWmedcost - RWfusionc)
  print("end")
}
 write.csv(RWvolbycountry, file="C:/Users/mubarsa/jrcbox/Forbiomod/SourceData/EFDM/data/costSupply/compareS2B_FUS_RW.csv")
 write.csv(PRvolbycountry, file="C:/Users/mubarsa/jrcbox/Forbiomod/SourceData/EFDM/data/costSupply/compareS2B_FUS_PR.csv")


```

